---
title: "IP-IgX correlations & WGCNA"
author: 
  - name: "Simon Couvreur"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    collapsed: false
    code_folding: hide
    number_sections: true
    self_contained: yes
    df_print: paged
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}

# run chunk to setup script for interactive use


# source("src/imports.R")
library(here)

figdir <- here("results/figures")
resultdir <- here("results")
cluster_pars_dir <- here("WGCNA_parameters")
datadir <- here("data")

knitr::opts_chunk$set(echo = TRUE,    # in combination with code_folding: hide
                      dev=c("png", "pdf"),
                      fig.path = here("results/figures/report/"),
                      out.width = '100%',
                      rows.print=30,   # options with df_print YAML setting https://bookdown.org/yihui/rmarkdown/html-document.html
                      cols.min.print=5)
# automatically switch .png to .pdf in knitr::include_graphics, when rendering to PDF
# https://bookdown.org/yihui/bookdown/figures.html
options(knitr.graphics.auto_pdf = TRUE) 
# vennDiagram uses futile.logger https://stackoverflow.com/questions/33393962/r-dont-write-log-file-for-venndiagram
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")


```

## Load data

```{r loadd}

loadd(ip_select_scores)
loadd(glycan_select_scores)
loadd(glycan_select_module_anno)
loadd(glycans_trans_fam_adj)
loadd(glycans_raw_omfo)
loadd(glycans_scale_free)
loadd(glycans_scale_free_noderiv)
loadd(ip_modules_df)
loadd(glycan_modules_df_noderiv)
loadd(ip_select_module_anno)
loadd(ips_trans_fam_adj)
loadd(ip_anno)
loadd(igx_mod_ips_n_associated_m)
loadd(glycan_modules_cor_ips_anno)
loadd(glycan_modules_cor_ips_modules_anno)
loadd(glycan_modality)
loadd(derived_glycan_pos)
loadd(IP_mod_varExpl_plots)
loadd(glycans_mod_varExpl_plots)
loadd(glycans_raw)
loadd(ip_select_modules)
```


# Within-modality correlation structures

## IPs

### Between modules

Large majority show little correlation, but some module pairs are extensively correlated.

```{r}

ip_module_cor <- cor(dplyr::select(ip_select_scores,-grey), method = "spearman")
ip_module_cor[upper.tri(ip_module_cor)] %>% 
  tibble(cor=.) %>% 
  ggplot(aes(x=cor))+
  geom_density(fill="blue", alpha=0.3)

```


## Glycans

### Between features

```{r}
corrplot::corrplot(cor(glycans_raw[, -(1:5)], use = "p"))
glycans_raw
```


### Between modules

```{r}

glycan_select_scores %>% 
  dplyr::select(-grey) %>% 
  set_names(glycan_select_module_anno$module_tag[match(names(.), 
                                                       glycan_select_module_anno$glycan_module)]) %>% 
  # dplyr::select(1:2) %>% 
  GGally::ggpairs(.,
                  lower=list(continuous=wrap("points", size=0.3)))

```

```{r}
hm_fontsize <- 7
glycan_module_cor <- cor(glycan_select_scores, method = "spearman")
corcol <-define_colors(as.vector(glycan_module_cor), breaks = c(-1,0,1), quant_colors = c("blue", "white", "red"))

cor_signed <- (glycan_module_cor+1)/2
cor_hc <- fastcluster::hclust(d=as.dist(1-cor_signed))
      
anno_hm <- rowAnnotation(
  module_tag=anno_text(glycan_select_module_anno$module_tag, gp = gpar(fontsize=hm_fontsize)),
  site=anno_text(glycan_select_module_anno$site, gp = gpar(fontsize=hm_fontsize)),
  chain=anno_text(glycan_select_module_anno$chain, gp = gpar(fontsize=hm_fontsize)),
  show_annotation_name=rep(TRUE,3)
  )
      
modality_df <- data.frame(feat_modality=glycan_select_module_anno$igx,
                          stringsAsFactors = FALSE)
modality_hm <- HeatmapAnnotation(df=modality_df,
                                 which = "row",
                                 show_annotation_name=FALSE,
                                 show_legend = TRUE,
                                 col=map(as_tibble(modality_df), ~define_colors(.x)),
                                 gp = gpar(col = "white", lwd = 0.5))

hm <- ComplexHeatmap::Heatmap(
  matrix=glycan_module_cor,
  name="spearman corr",
  col= corcol,
  cluster_rows = as.dendrogram(cor_hc),
  cluster_columns = rev(as.dendrogram(cor_hc)),
  row_names_gp = gpar(fontsize = 12),
  column_dend_side = "top",
  show_column_names = FALSE,
  right_annotation = modality_hm,
  cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(sprintf("%.1f", glycan_module_cor[i, j]), x, y, gp = gpar(fontsize = hm_fontsize))
  },
  row_labels = glycan_select_module_anno$module_tag %>% na.omit()
)

draw(hm+anno_hm, heatmap_legend_side = "top", annotation_legend_side = "top",
     legend_labels_gp=gpar(fontsize = hm_fontsize), 
     legend_title_gp=gpar(fontsize = hm_fontsize))

```

# WGCNA

Objective of WGCNA is not necessarily globally optimal partitioning, but rather detection of meaningful modules. Grey module is allowed to be very heterogeneous. Also, scale-free property is mostly a heuristic to downscale noise via the power transform, rather than based on theoretical arguments from network/graph theory. Ref Peter Langfelder, Bioconductor forum, https://support.bioconductor.org/p/66101/ [24/04/2019].

How to deal with "grey" module?
  
  Current approach:
  - treat as other clusters, but ignore module-specific modularity contribution in calculation of global modularity?
  - Equivalently: treat all grey members as singleton clusters (so that in modularity calculation: delta(C_i,C_j)=0 for every singleton cluster i and every cluster j)

But also included full modularity, treating the grey cluster as a real cluster.

Modularity is calculated based on three similarity matrices:
  - correlation (following Gomez2009, with known limitations cf MacMahon2015).
- adjacency ~ network parameters
- TOM ~ network parameters

## Glycans

### Excluding derived traits

#### General remarks

Negative correlations can carry important information, especially when the glycans are separated by addition/removal of a single sugar moiety.

To assure this

1) Structure of the glycan would need to be known
2) Biological evidence of the reaction, e.g. compiled in database such as KEGG?

But: sialic acid typically terminal -> excellent candidate

Examples: 
ENI1H5N4F0S1 and ENI1H5N4F0S2 are strongly negatively correlated.
LAGY1H5N4F0S2 and LAGY1H5N4F0S1 are strongly positively correlated.

Could relate to:

* regulation of reaction: constitutive versus induced?
* opposite associations with IPs?

Some derived traits might be driven primarily by a single glycan. Example: ENI_A2S and ENI1H5N4F0S2

-> Possible to assess 'relevance' of derived traits by variance decomposition?

#### power parameter

```{r glycans-scale-free-softpower-noderiv}

# View(glycans_scale_free$sft_df)
cowplot::plot_grid(
  glycans_scale_free_noderiv$p_rsq,
  glycans_scale_free_noderiv$p_k
)
#low rsq with power distribution (scale free topology): because not independent entities but related measurements?
```



#### Optimal cut
Criteria per cut:
  Distributions (per cut, per module) of modularity and varExplained do not track.
  
VarExplained-constrained (treshold 0.5) modularity:

```{r, fig.asp=2.5}
glycans_mod_varExpl_plots$p_mod_sum_varexpl_constraint
```

VarExplained-weighted modularity:

```{r, fig.asp=2.5}
glycans_mod_varExpl_plots$p_mod_sum_varexpl_weighted
```

##### Cut selection

View corrsponding heatmap (selected cut = 3, third column in left-hand panel)

```{r eval=FALSE, fig.asp=2, include=FALSE}

p=glycan_modules_df_noderiv %>% 
  dplyr::filter(
    corfnc=="bicor",
    networktype=="signed_hybrid",
    powers==6,
  ) %>% 
  dplyr::select(plots) %>% 
  unlist()
draw(p$plots, heatmap_legend_side = "top", annotation_legend_side = "top")

# # feature cluster labels
# glycan_modules_df_noderiv %>%
#   unnest(cut_stats) %>%
#   dplyr::filter(
#     corfnc=="bicor",
#     networktype=="signed_hybrid",
#     powers==6,
#     cut==3
#   ) %>%
#   unnest(feature_stats) %>%
#   group_by(colors) %>% 
#   mutate(n=n()) %>% 
#   View()

```

Clean heatmap, incl extra glycans annotations

```{r fig.asp=2}

row_names_fontsize <- 1
legend_fontsize <- 15 

corr_like <- WGCNA::bicor(x=glycans_trans_fam_adj[,!derived_glycan_pos], use = "pairwise.complete.obs")
similarity <- similarity_from_cor(cordat = corr_like, networktype = "signed_hybrid")
adj <- as.matrix(similarity)^6
TOM <- TOMsimilarity(adj)
dissTOM <-1 - TOM
feat_hc_TOM <- fastcluster::hclust(as.dist(dissTOM), method = "average")
color_overview <- glycan_modules_df_noderiv %>% 
  unnest(cut_stats) %>% 
  unnest(feature_stats) %>% 
  dplyr::filter(
    corfnc=="bicor",
    networktype=="signed_hybrid",
    powers==6,
    cut==3
  )

# dfs for additional annotation
modality_df <- data.frame(
    linkage = ifelse(str_detect(color_overview$feat, pattern = "HYT"),
                     "O-glycan",
                     "N-glycan"),
    `immunoglobulin`=glycan_modality[!derived_glycan_pos],
    `anchor site`=color_overview$feat %>% 
      str_split(pattern = "1H") %>% 
      map_chr(1) %>% 
      ifelse(str_detect(., pattern = "IgG"), ., paste0("IgA-",.)),
    check.names = FALSE
  )
glycan_composition <- map(structure(color_overview$feat,names=color_overview$feat),
    ~map2(structure(c("H(\\d)","N(\\d)","F(\\d)","S(\\d)"),
                    names=c("hexose", "N-acetylhexoseamine", "fucose", "sialic acid")),
         .x,
        function(rgx, feat){
          str_match(string=feat, pattern = rgx)[,2] %>% 
            replace_na(0) %>% 
            as.numeric()
        }) %>% 
      bind_cols()
    ) %>% 
  bind_rows(.id="feat") %>% 
  mutate(fucose=as.character(fucose))

# manual color annoation to set categorical label for fucose (0/1)
glycan_composition_cols <- list(
  hexose = define_colors(glycan_composition$hexose, breaks = c(0,6)),
  `N-acetylhexoseamine` = define_colors(glycan_composition$`N-acetylhexoseamine`, breaks = c(0,6)),
  fucose = c("0"="white", "1"="darkgreen"),
  `sialic acid` = define_colors(glycan_composition$`sialic acid`, breaks = c(0,6))
)

# heatmaps for additional annotations
hm_anno_glycan_comp <- rowAnnotation(df=as.data.frame(glycan_composition[,-1]),
                                     col=glycan_composition_cols)
hm_feat_text <- rowAnnotation(module_text=anno_text(color_overview$feat),
                               gp = gpar(fontsize=row_names_fontsize))

color_overview_by_module_tag <- data.frame(
  module=glycan_select_module_anno$module_tag[
    match(color_overview$colors,glycan_select_module_anno$glycan_module)] %>% 
      replace_na("grey"))

hm_module_text <- rowAnnotation(module_text=anno_text(color_overview_by_module_tag$module %>% 
                                                     str_remove(pattern = "grey")),
                             gp = gpar(fontsize=row_names_fontsize))

# call function to build core heatmap + 'standard' annotations
hml <- make_heatmap_list(
  corr_like = corr_like,
  feat_hc_TOM=feat_hc_TOM,
  color_overview = color_overview_by_module_tag,
  modality_df = modality_df,
  row_names_fontsize = row_names_fontsize,
  heatmap_body_legend_title = "bicor"
)

#draw result
hml2 <- draw(hm_module_text + hml+hm_feat_text+hm_anno_glycan_comp, heatmap_legend_side = "top", annotation_legend_side = "top",
     legend_labels_gp=gpar(fontsize = legend_fontsize), 
     legend_title_gp=gpar(fontsize = legend_fontsize))
# draw(hml2)

# add module separators
annotations_to_separate <- list_components() %>%
  str_match(.,pattern = "^annotation_(.*)_1$") %>%
  .[,2] %>% 
  na.omit() %>% 
  unique() %>% 
  setdiff(., "module")


module_transitions <- c()
output_modules <- color_overview_by_module_tag$module[feat_hc_TOM$order]
for (i in seq_len(length(output_modules)-1)){
  module_transitions[i] <- output_modules[i] != output_modules[i+1]
}
module_transition_pos <- which(rev(module_transitions))/(length(module_transitions)+1)
walk(module_transition_pos, 
     .f = function(x){
       
       # horizontal lines in heatmap body
       decorate_heatmap_body(
         heatmap = "bicor",
         code={grid.lines(c(0, 1), c(x, x), gp = gpar(lty = 2, lwd = row_names_fontsize/20))}
       )
       
       # horizontal lines in annotation columns
       walk(annotations_to_separate, 
            function(anno){
              decorate_annotation(
                annotation=anno,
                code={grid.lines(c(0, 1), c(x, x), gp = gpar(lty = 2, lwd = row_names_fontsize/20))}
              )
            }
       )
     }
)


```


  
## Immunophenotypes
  
### Optimal cut

Cuts with deep==4 (i.e. uneven cut indices) perform consistently worse on Gomez modularity.

Ranking is very similar (and identical top 3) when either varExplained-thesholding of varExplained-weighting is used.

VarExplained-constrained (treshold 0.5) modularity:

```{r, fig.asp=2.5}
IP_mod_varExpl_plots$p_mod_sum_varexpl_constraint
```

VarExplained-weighted modularity:

```{r, fig.asp=2.5}
IP_mod_varExpl_plots$p_mod_sum_varexpl_weighted
```


### Cut selection

Explore selected cut

```{r}

ip_cut <- ip_modules_df %>% 
  unnest(cut_stats) %>% 
  dplyr::filter(
    corfnc=="spear",
    networktype=="signed_hybrid",
    powers==5,
    cut==8
  )

# feature labels
# ip_cut %>% 
#   unnest(feature_stats) %>% 
#   View()

```

Number of modules: 585

```{r}
ip_cut %>% 
  unnest(feature_stats) %>% 
  summarise(n=n_distinct(colors)) %>% 
  .$n
```



Module size distribution (excluding grey cluster).

```{r}
ip_cut%>% 
  unnest(module_stats) %>% 
  dplyr::filter(colors!="grey") %>% 
  .$n %>% 
  boxplot()

```


#### Create annotations

Output to link glycans annotations with glycan names in heatmap:

```{r}
# glycan_modules_df_noderiv %>% 
#   unnest(cut_stats) %>% 
#   unnest(feature_stats) %>% 
#   dplyr::filter(
#     corfnc=="bicor",
#     networktype=="signed_hybrid",
#     powers==6,
#     cut==3
#   ) %>% View()



```


#### Explore annotations

Modules are highly homogenous w.r.t. source. 

Cumulative distribution of the proportion, per module, of features in the module
which belong to the most abundant IP source (i.e. subclass such as CD8) in that module.

```{r}

plot(ecdf(ip_select_module_anno$max_prop),
     xlab="Proportion of features (in module) which belong to most abundant IP source",
     ylab="Cumulative proportion of modules")
  
```

number and size of modules per source

```{r}

ip_select_module_anno %>%
  ggplot(aes(x=label, y=module_size))+
  geom_boxplot(varwidth = TRUE)+
  coord_flip()

```


Distribution of module sizes

```{r}
histogram(ip_select_module_anno$module_size, nint=diff(range(ip_select_module_anno$module_size, na.rm=TRUE)))

power2 <- 2^(1:10)
span_power_2 <- matrix(power2, ncol = 1) %*% matrix(1+((0:5)*2),nrow = 1) %>% as.vector() %>% .[.<600] 
span_power_2_only <- setdiff(span_power_2, power2)

tibble(size =ip_select_module_anno$module_size,
       power2 = size %in% power2,
       subspan_power2_only = size %in% span_power_2_only,
       power_anno = ifelse(power2, "power2=2^x, x <=10", 
                           ifelse(subspan_power2_only, "subspan_power2= a*power2, a uneven <=11", "other"))) %>% 
# tibble(size =ip_select_module_anno$module_size,
#        power2 = size %in% power2,
#        power_anno = ifelse(power2, "power2=2^x", 
#                            ifelse(size %in% 128*c(3,5), "span_power_2= a*2^x, a uneven <=11",
#                                   ifelse (size %in% 64 )))) %>% 
  drop_na() %>% 
  ggplot(aes(x=size, fill=power_anno))+
  geom_histogram(binwidth=1)

```


##### IP subclass representation

compare module-source distribution with feature-source distribution. 

(here the source of a module is the most frequent source in the module ('label_major'),
which is an appropriate approximation because modules are typically highly homogeneous
wrt source, see above)

Representation depends on number of features per source:

- sources with few features -> no representation
- sources with many features -> representation proportional to number of features per source

```{r}

full_join(
  count(ip_select_module_anno, label_major) %>% rename("composite_lin_source"="label_major"),
  dplyr::filter(ip_anno, set_name %in% names(ips_trans_fam_adj)) %>% 
    dplyr::count(composite_lin_source),
  by=c("composite_lin_source"),
  suffix=c("_module", "_feature")
) %>%
  dplyr::filter(!is.na(composite_lin_source)) %>% 
  mutate(n_module=replace_na(n_module,0)) %>% 
  arrange(n_feature) %>% 
  mutate(composite_lin_source=factor(composite_lin_source,composite_lin_source)) %>% 
  tidyr::gather(dataset, n, -composite_lin_source) %>% 
  ggplot(aes(x=reorder(composite_lin_source,n), y=n))+
  geom_bar(stat = "identity")+
  coord_flip()+
  facet_wrap(~dataset, scales="free_x")

```


As a results: overrepresentation in the grey module:

```{r}
grey_represent <- ip_modules_df %>% 
    unnest(cut_stats) %>% 
    dplyr::filter(
      corfnc=="spear",
      networktype=="signed_hybrid",
      powers==5,
      cut==8
    ) %>% 
    unnest(feature_stats) %>% 
    left_join(ip_anno %>% 
                dplyr::select(set_name, subset_name, composite_lin_source), 
              by=c("feat"="set_name")) %>% 
    count(colors, composite_lin_source)%>% 
  dplyr::filter(colors=="grey") %>% 
  full_join(
    dplyr::filter(ip_anno, set_name %in% names(ips_trans_fam_adj)) %>% 
      dplyr::count(composite_lin_source),
    by="composite_lin_source",
    suffix=c("grey", "overall")) %>% 
  mutate(prop_grey=ngrey/noverall) %>% 
  arrange(desc(prop_grey)) 

grey_represent%>% 
  ggplot(aes(x=noverall, y=prop_grey, color=composite_lin_source))+
  geom_point()+
  scale_color_discrete(guide="none")+
  geom_text_repel(aes(label=composite_lin_source))+
  theme_bw()+
  facet_zoom(xlim=c(0,4000))+
  geom_vline(xintercept = 500, color="red")+
  xlab("Number of IPs in subclass")+
  ylab("Proportion in grey module")
```

The correlation structure can differ substantially between subclasses (parent lineages). Specifically,
as more markers are available for a parent lineage,
the number of features for that lineage increases exponentially (2^markers).
One single immune cell can contribute towards multiple CSFs of the same
parent lineage, but of different resolution (e.g. a cell of lineage X characterised 
by markers A, B and C [A+B+C+] would contribute to the CSFs A+, B+, C+,
A+B+, B+C+, A+C+ and A+B+C+ of lineage X). Therefore
within-subclass feature correlations can be expected to be larger on average than 
between-subclass correlations. 

The combined result of these two is that
subclasses with larger number of features (i.e. which are phenotyped deeper, with more markers)
can dominate both the parameter tuning (because parameters are tuned to the highly-abundant, relatively stronger correlated features of deeply phenotyped lineages) and module detection stages (hierarchical clustering and dendrogram cutting). In addition the modularity criterion applied in the module detection heuristic
is itself subject to a resolution limit, whereby the criterion is insensitive to 
smaller clusters.

The overall results is that a substantial proportion of features in small IP subclasses remaining unassigned, and thus excluded from subsequent analyses if those analyses are
based only on the defined modules.

- The fact that the problem arises specifically for groups with a small number of features
allows to analyse univariate associations with glycan modules, without substantial increase in the multiple testing burden.
- Separate WGCNA on small subclasses could be performed
- Integrate multiscale approaches to module definition e.g. MEDENA (Song2015)

In this case 83.6% of features in small IP subclasses (defined arbitrarily as <500 features)
are unassigned. As they make up only 2% of all features, a univariate analysis
association analysis with glycans modules could be considered.


```{r}
summ <- grey_represent %>%
  group_by(noverall>500) %>% 
  summarise(ngrey_tot = sum(ngrey),
            noverall_tot = sum(noverall),
            perc_grey = round(ngrey_tot/noverall_tot*100, 1)) 

summ
summ[1,2:3]/apply(summ[,2:3],2,sum)    
```

correlation density plot for large vs small IP sources

```{r}

```




# Cross-Modality correlations

## Glycan modules - IPs

Batch effects of lesser relevance.

### Number of significantly associated features per feature group ~ FDR threshold


```{r n-sign-ass-features-FDR}



igx_mod_ips_n_associated_m %>% 
  ungroup() %>%
  dplyr::rename("feature_group"="omic") %>% 
  mutate(
    feature_group = dplyr::recode(
      feature_group,
      IgA="IgA_module",
      IgG="IgG_module",
      modules="IgX_module")
  ) %>% 
  ggplot(aes(x=fdr_threshold, y=prop, colour=feature_group))+
  geom_step()+
  theme_bw()+
  # facet_zoom(x = fdr_threshold < 0.25)
  facet_zoom(xlim = c(0,0.2))+
  xlab("q value")+
  ylab("proportion of features with association < FDR")+
  geom_vline(xintercept=0.05, color="grey")

```


### IP subgroup enrichment

Problem with all IP subgroup enrichments (incl enrichment stratified by module,
incl when subsequently pooled by annotation variable) is that IPs are
intercorrelated, causing inflation of the variance. See e.g. ?limma::camera (Wu & Smyth 2012)

#### Among q (storey) < 0.05 (dichotomous)

Overall

```{r}

glycan_modules_cor_ips_anno %>% 
  # dplyr::mutate(storey_sign = factor(ifelse(p.val_storey<0.05, "ip_sign", "ip_nonsign"))) %>% 
  group_by(composite_lin_source) %>% 
  summarise(
    n_tests_source = n(),
    n_tests = nrow(glycan_modules_cor_ips_anno),
    n_sign_ip = sum(p.val_storey<0.05),
    n_sign_test = sum(glycan_modules_cor_ips_anno$p.val_storey<0.05),
  ) %>%
  ungroup %>% 
  mutate(p_hyper = ifelse(
    n_sign_ip>0,
    phyper(q=n_sign_ip, m=n_tests_source,
                     n=n_tests-n_tests_source, k=n_sign_test,
                     lower.tail = FALSE),
    1)
  ) %>% 
  arrange(p_hyper)
  
  
  
```



```{r}

cor_enrich_by_group <- function(cor_anno, anno_group, enrich_group, bool_sig_col){
  # enrichment (competitive, Goeman2007) of levels of enrich_group, stratified by anno_group
  
  anno_group <- enquo(anno_group)
  bool_sig_col <- enquo(bool_sig_col)
  enrich_group <- enquo(enrich_group)
  
  cor_anno %>% 
    # dplyr::mutate(storey_sign = factor(ifelse(p.val_storey<0.05, "ip_sign", "ip_nonsign"))) %>% 
    group_by(!!enrich_group, !!anno_group) %>% 
    summarise(
      n_tests_enrich_group = n(),
      n_sign_enrich_group = sum(!!bool_sig_col),
    ) %>%
    group_by(!!anno_group) %>% 
    mutate(
      n_tests_anno_group = sum(n_tests_enrich_group),
      n_sign_anno_group = sum(n_sign_enrich_group),
    ) %>% 
    ungroup() %>% 
    # dplyr::filter(n_sign_anno_group>1) %>%    # remove anno_groups where denominator is 1?
    mutate(
      p_hyper = ifelse(
        n_sign_enrich_group>0,
        phyper(q=n_sign_enrich_group, m=n_tests_enrich_group,
               n=n_tests_anno_group-n_tests_enrich_group, k=n_sign_anno_group,
               lower.tail = FALSE),
        1),
      q_hyper = qvalue(p_hyper, pi0.method="smoother")$qvalues
    ) %>% 
    dplyr::filter(q_hyper < 0.05) %>% 
    arrange(!!anno_group, p_hyper) %>% 
    dplyr::select(!!anno_group, !!enrich_group, p_hyper, q_hyper, everything())
}
  
```



Stratified by glycan module

```{r}

glycan_modules_cor_ips_anno %>% 
  mutate(bool_sig_col = p.val_storey < 0.05) %>% 
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = module_tag, enrich_group=composite_lin_source) %>% 
  left_join(glycan_select_module_anno)

```

Glycan modules can be grouped by annotation, and the enrichment calculated among
all tests which are run for each group of the annotation variable (i.e. pooling
the number of tests and the number of significant tests per group).

Stratified by IgX

```{r}

glycan_modules_cor_ips_anno %>% 
  mutate(bool_sig_col = p.val_storey < 0.05) %>% 
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = igx, enrich_group=composite_lin_source)
  
```

Stratified by fucosylation

```{r}

glycan_modules_cor_ips_anno %>% 
  mutate(bool_sig_col = p.val_storey < 0.05) %>% 
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = fucose, enrich_group=composite_lin_source)

```

Stratified by sialylation

```{r}

glycan_modules_cor_ips_anno %>% 
  mutate(bool_sig_col = p.val_storey < 0.05) %>% 
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = sialic_acid, enrich_group=composite_lin_source)

```


## Glycan modules - IP modules

### Top associations

```{r}
glycan_modules_cor_ips_modules_anno %>%
  dplyr::filter(IP_module != "grey" & p.val_storey<0.05)
# glycan_modules_cor_ips_modules_anno %>% 
#   dplyr::filter(IP_module != "grey" & p.val<0.05)

```

```{r}
top_mod_mod <- glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(p.val_storey < 0.05) %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(core_CD = map(IP_feature_core_data, ~unique(.x$core_CD))) %>% 
  dplyr::select(module_tag_glycan, igx,site, fucose, sialic_acid, hexose,
                module_tag_IP,  label_perc, label_major, core_CD,
                p.val_storey, correlation, varExplained_IP,
                IP_feature_core_data, IP_module) 

```

#### Correlation structure between top associations

```{r}

top_mod_hex <- dplyr::select(top_mod_mod,IP_module)

ip_select_scores[, unlist(top_mod_hex)] %>% 
  set_names(
    glycan_modules_cor_ips_modules_anno$module_tag_IP[
      match(names(.), glycan_modules_cor_ips_modules_anno$IP_module)]
  ) %>% 
  cor(.) %>% 
Heatmap(.)

```

Clean heatmap, incl extra glycans annotations

```{r fig.asp=2}

row_names_fontsize <- 10
legend_fontsize <- 35
glycan_module_site_hm_cm <- 3
ip_anno_hm_cm <- 1
ip_lineage_anno_cm<-3
ip_core_hm_width=7
core_hm_width=25

glycan_module_anno_df <- top_mod_mod %>% 
  mutate(Ig_site = recode(site, "N-glycan"="IgA"))

corr_like_l <- ip_select_scores[, unlist(top_mod_hex)] %>% 
  set_names(
    glycan_modules_cor_ips_modules_anno$module_tag_IP[
      match(names(.), glycan_modules_cor_ips_modules_anno$IP_module)]
  ) %>% 
  WGCNA::corAndPvalue(., use = "pairwise.complete.obs")p

# calculate core heatmap first, so row order can be used in annotation heatmaps
# call function to build core heatmap + 'standard' annotations
corcol <-define_colors(as.vector(corr_like_l$cor), breaks = c(-1,0,1), quant_colors = c("blue", "white", "red"))
core_hm <- ComplexHeatmap::Heatmap(
  matrix=corr_like_l$cor,
  name="IP-IP correlation (pearson)",
  col= corcol,
  cluster_rows = TRUE,
  cluster_columns = TRUE,
  show_row_dend = FALSE,
  row_names_gp = gpar(fontsize = row_names_fontsize),
  column_dend_side = "top",
  # right_annotation = modality_hm,
  cell_fun = function(j, i, x, y, width, height, fill) {
    pv <-corr_like_l$p[i,j]
    star <- ifelse(
      pv < 0.005, "***",
      ifelse(pv <0.01, "**",
             ifelse(pv<0.05, "*","")
      )
    )
    grid.text(sprintf("%s",star), x, y, gp = gpar(fontsize = row_names_fontsize))
  },
  split=glycan_module_anno_df$module_tag_glycan,
  column_split = glycan_module_anno_df$module_tag_glycan,
  cluster_row_slices = FALSE,
  cluster_column_slices = FALSE,
  width=core_hm_width,
  row_title = NULL,
  show_column_names = TRUE,
  column_names_side = "bottom"
)


glycan_comp_anno_hm <- rowAnnotation(
  df=dplyr::select(top_mod_mod, fucose, sialic_acid, hexose) %>% 
    mutate_all(~recode(.x, low="low/none",
                      none="low/none")) %>% 
    mutate_all(~factor(.x, 
                       levels=c("low/none","average/mixed", "all"))),
  col=list(
    fucose=c(`low/none`="#f4cdcd", `average/mixed`="white", all="#bdf2b8"),
    sialic_acid=c(`low/none`="#f4cdcd", `average/mixed`="white", all="#bdf2b8"),
    hexose=c(`low/none`="#f4cdcd", `average/mixed`="white", all="#bdf2b8")
  )
    # fucose=c(none="white", `average/mixed`="lightgreen", all="darkgreen"),
    # sialic_acid=c(low="white", `average/mixed`="lightgreen"),
    # hexose=c(low="white", `average/mixed`="lightgreen")
)

#works because in this case glycan tag <-1:1-> glycan site (i.e. no two tags from same site)
glycan_module_site_hm <- Heatmap(
  as.matrix(glycan_module_anno_df$Ig_site),
  col = c(IgGI="#a4a8a4", IgGII="#727572", `IgGI/II/IV`="#4e4f4e", IgA="#bee0e8"),
  cell_fun = function(j, i, x, y, width, height, fill) {
    site <-glycan_module_anno_df$Ig_site[i]
    glycan_module <-glycan_module_anno_df$module_tag_glycan[i]
    module_association_line_ranks <- row_order(core_hm)[[glycan_module]]
    association_line_rank <- which(module_association_line_ranks==i)
    if (association_line_rank == ceiling(length(module_association_line_ranks)/2))  #print only the line in the middle of the module
      grid.text(sprintf("%s (%s)",glycan_module, site), x, y, gp = gpar(fontsize = row_names_fontsize))
  },
  show_heatmap_legend = FALSE,
  width=unit(glycan_module_site_hm_cm,"cm")
)

# glycan module - IP module correlations
corcol_mod_cor <-define_colors(as.vector(top_mod_mod$correlation), breaks = c(-1,0,1), quant_colors = c("blue", "white", "red"))
hm_cor_mod <- Heatmap(
  name="glycan-IP correlation (pearson)",
  as.matrix(ifelse(top_mod_mod$correlation>0, "pos", "neg")),
  col = c(pos="#b25e5e", neg="#3a4d82"),
  cell_fun = function(j, i, x, y, width, height, fill) {
    pv <-top_mod_mod$p.val_storey[i]
    star <- ifelse(
      pv < 0.005, "***",
      ifelse(pv <0.01, "**",
             ifelse(pv<0.05, "*","")
      )
    )
    grid.text(sprintf("%s",star), x, y, gp = gpar(fontsize = row_names_fontsize))
  }
)


ip_anno_hm<- rowAnnotation(
  module_tag_IP=anno_text(top_mod_mod$module_tag_IP, gp = gpar(fontsize=row_names_fontsize)),
    width=unit(ip_anno_hm_cm,"cm")
)


ip_lineage_anno <-  Heatmap(
  as.matrix(top_mod_mod$label_major),
  col = c(CD8="#A8A8A8", DPT="#999999", Monocytes="#898989", `CD4/Naive`="#797979", `B cells/G`="#686868"),
  cell_fun = function(j, i, x, y, width, height, fill) {
    site <-top_mod_mod$label_major[i]
    grid.text(sprintf("%s",site), x, y, gp = gpar(fontsize = row_names_fontsize))
  },
  show_heatmap_legend = FALSE,
  width=unit(ip_lineage_anno_cm,"cm")
)


ip_CD_core_equiv <- pmap_chr(
  dplyr::select(top_mod_mod, label_major, IP_feature_core_data), 
  function(label_major, IP_feature_core_data)     # extract core_CD for major label of each module
    IP_feature_core_data[IP_feature_core_data$composite_lin_source==label_major,
               "core_CD_equiv",
               drop=TRUE] %>% 
    unique()
)

ip_CD_core_spread <- ip_CD_core_equiv  %>% 
  str_match_all(pattern="([^(/+\\-)]+)([+\\-]{1})") %>%  #spread out
  set_names(top_mod_mod$module_tag_IP) %>%
  map(data.table) %>%
  bind_rows(.id="module_tag_IP") %>%
  set_names(c("module_tag_IP", "match", "CD", "status")) %>%
  dplyr::select(-match) %>%
  spread(CD, status) %>% 
  left_join(top_mod_mod[,'module_tag_IP', drop=FALSE],.,by="module_tag_IP")  # same order as top_mod_mod

ip_CD_core_spread_shared <- ip_CD_core_spread %>%  # markers shared between at least two IP modules
  dplyr::select(-module_tag_IP) %>%
  dplyr::select_if(~sum(!is.na(.x))>1) %>% 
  mutate_all(function(x)replace_na(x,"na"))

ip_core_text_hm <- rowAnnotation(
  core_equiv = anno_text(ip_CD_core_equiv,gp = gpar(fontsize=row_names_fontsize))
)

ip_core_hm <- Heatmap(
  as.matrix(ip_CD_core_spread_shared),
  col=c(`+`="#bdf2b8", `-`="#f4cdcd", na="white"),
  cell_fun = function(j, i, x, y, width, height, fill) {
    state <-ip_CD_core_spread_shared[i,j]
    if (!state=="na")
      grid.text(sprintf("%s",state), x, y, gp = gpar(fontsize = row_names_fontsize))
  },
  width = ip_core_hm_width,
  show_heatmap_legend = FALSE
)



draw(glycan_comp_anno_hm+glycan_module_site_hm+
       hm_cor_mod+
       core_hm+
       ip_anno_hm+ip_lineage_anno+ip_core_text_hm+ip_core_hm,
     heatmap_legend_side = "top", annotation_legend_side = "top",
     legend_labels_gp=gpar(fontsize = legend_fontsize), 
     legend_title_gp=gpar(fontsize = legend_fontsize),
     main_heatmap=4, merge_legends=TRUE)

list_components()


```

### Enrichment

Use in report: 

1) nominal pv threshold -> hypergeometric
2) global (i.e. entire list) -> KS

#### Dichotomised

##### Enrichment threshold = adjusted (storey)

**Generally not meaningful due to small overall number of significant associations.**


```{r}

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val_storey < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = module_tag_glycan, enrich_group = label_major) %>% 
  left_join(glycan_select_module_anno, by=c("module_tag_glycan"="module_tag"))

```

Glycan modules grouped by igx

```{r}

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val_storey < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = igx, enrich_group = label_major)

```

Glycan modules grouped by fucose

```{r}

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val_storey < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = fucose, enrich_group = label_major)

```

Glycan modules grouped by sialylation

```{r}

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val_storey < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = sialic_acid, enrich_group = label_major)

```


##### Enrichment threshold = nominal

```{r}
glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey" & p.val<0.05) %>% 
  dim
```

For each glycan module:

```{r}

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = module_tag_glycan, enrich_group = label_major) %>% 
  left_join(glycan_select_module_anno, by=c("module_tag_glycan"="module_tag"))

```

Glycan modules grouped by igx

```{r}

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = igx, enrich_group = label_major)

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = label_major, enrich_group = igx)

```

Glycan modules grouped by fucose

```{r}

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = fucose, enrich_group = label_major)

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = label_major, enrich_group = fucose)

```

Glycan modules grouped by sialylation

strongest enrichment for CD8 - low sialyl (both ways)

```{r}

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = sialic_acid, enrich_group = label_major)

glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  mutate(bool_sig_col = p.val < 0.05) %>%   # among nominally significant tests
  cor_enrich_by_group(bool_sig_col = bool_sig_col, anno_group = label_major, enrich_group = sialic_acid)
```

#### Rank-based (global or pv-truncated)

MWW never gives significant enrichment.

KS gives:

Among nominally significant p-values:
- only marginal enrichment towards the top (at FDR < 0.05) is for glycan module G1
- none of the stratified analyses (by igx, fucose or sialic acid) show enrichment towards the top at FDR < 0.05.

(At p-value threshold 0.1: same as at 0.05, and

- IP modules of source 'CD4/Naive' significantly enriched towards the top among associations with glycan modules of mixed fucose content. )

In global analysis (i.e. over full pv-ranked association list):

- Marginal enrichment for glycan modules G1 and G5
- Among correlations with IgG modules: IP modules of source 'Monocytes' are enriched towards the top.
- Among correlations with fully fucosylated glycan modules: IP modules of source 'Monocytes' are enriched towards the top.
- Among correlations with modules containing glycans of mixed/average sialylation levels: IP modules of sources 'CD4/Naive', 'CD4', '16+56-', 16+56' and 'B cells/MD+' are enriched towards the top.
- Among correlations with modules containing glycans of low sialylation levels: IP modules of sources 'CD8' and 'Monocytes' are enriched towards the top.


```{r}
calc_ks <- function(x, enrich_group){
  
  p_ks <- c()
  for (label in unique(x[,enrich_group])){
    p_ks[label] <- ks.test(
      x=x$p.val[x[,enrich_group]==label],
      y=x$p.val[x[,enrich_group]!=label],
      alternative = "greater"  
      # cf ?ks.test: "in the two-sample case alternative = "greater" includes 
      # distributions for which x is stochastically smaller than y 
      # (the CDF of x lies above and hence to the left of that for y)"
    )$p.value
  }
  # browser()
  p_ks %>% 
    data.frame(p.val=.) %>% 
    rownames_to_column("enrich_group") %>% 
    mutate(q_storey = qvalue(p.val, pi0.method="smoother")$qvalues) %>% 
    dplyr::filter(p.val<0.05)
}

calc_mww <- function(x, enrich_group){
  
  p_list <- list()
  for (label in unique(x[,enrich_group])){
    ranks <- x$p.val[x[,enrich_group]==label]
    ranks_ref <- x$p.val[x[,enrich_group]!=label]
    
    p_list[[label]] <- list(
      p= wilcox.test(
        x=ranks,
        y=ranks_ref
      )$p.value,
      diff_median = median(ranks)-median(ranks_ref)
    )
  }
  p_list[p_list$p<0.05]
}

rank_based_by_group <- function(x, anno_group, enrich_group, bool_sig_col){
  bool_sig_col <- enquo(bool_sig_col)
  
  tmp <- x %>% 
  dplyr::filter(IP_module != "grey") %>% 
  arrange(p.val) %>% 
  mutate(rank=seq_len(nrow(.))/nrow(.)) %>% 
  # head(1000)
  dplyr::filter(bool_sig_col) %>%
  split(.[,anno_group])
  
  list(
    ks=map(tmp, calc_ks, enrich_group=enrich_group),
    mww=map(tmp, calc_mww, enrich_group=enrich_group)
  )
}

```

Parameter

```{r}
# pval_thres <- 0.05
# pval_thres <- 0.1
pval_thres <- 2



# Overall



l <- glycan_modules_cor_ips_modules_anno %>% 
  dplyr::filter(IP_module != "grey") %>% 
  arrange(p.val) %>% 
  dplyr::filter(p.val < pval_thres)

calc_ks(l, enrich_group="label_major")
calc_mww(l, enrich_group="label_major")
calc_ks(l, enrich_group="module_tag_glycan")
calc_mww(l, enrich_group="module_tag_glycan")




# By Igx


glycan_modules_cor_ips_modules_anno %>% 
  mutate(bool_sig_col = p.val<pval_thres) %>% 
rank_based_by_group(enrich_group = "label_major",
            anno_group = "igx",
            bool_sig_col=bool_sig_col)



# By fucose


glycan_modules_cor_ips_modules_anno %>% 
  mutate(bool_sig_col = p.val<pval_thres) %>% 
rank_based_by_group(enrich_group = "label_major",
            anno_group = "fucose",
            bool_sig_col=bool_sig_col)



# By sialic acid


glycan_modules_cor_ips_modules_anno %>% 
  mutate(bool_sig_col = p.val<pval_thres) %>% 
rank_based_by_group(enrich_group = "label_major",
            anno_group = "sialic_acid",
            bool_sig_col=bool_sig_col)

```



# Session info

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```


# References

















