---
  title: "IP-IgX correlations & WGCNA"
author: 
  - name: "Simon Couvreur"
output: 
  html_document:
  toc: true
toc_float: true
toc_depth: 5
collapsed: false
code_folding: hide
number_sections: true
self_contained: yes
editor_options: 
  chunk_output_type: console
---

# Setup
  
```{r setup, include=FALSE}

# run chunk to setup script for interactive use


# source("src/imports.R")

figdir <- here("results/figures")
resultdir <- here("results")
cluster_pars_dir <- here("WGCNA_parameters")
datadir <- here("data")

knitr::opts_chunk$set(echo = TRUE,    # in combination with code_folding: hide
                      dev=c("png", "pdf"),
                      fig.path = here("results/figures/report/"))
# automatically switch .png to .pdf in knitr::include_graphics, when rendering to PDF
# https://bookdown.org/yihui/bookdown/figures.html
options(knitr.graphics.auto_pdf = TRUE) 
# vennDiagram uses futile.logger https://stackoverflow.com/questions/33393962/r-dont-write-log-file-for-venndiagram
futile.logger::flog.threshold(futile.logger::ERROR, name = "VennDiagramLogger")


```

## Load data

```{r loadd}


```


# Within-modality correlation structures

## IPs

### Between modules

Large majority show little correlation, but some module pairs are extensively correlated.

```{r}

ip_module_cor <- cor(dplyr::select(ip_select_scores,-grey), method = "spearman")
ip_module_cor[upper.tri(ip_module_cor)] %>% boxplot(.)

```


## Glycans

### Between modules

```{r}

GGally::ggpairs(glycan_select_scores)

```

```{r}

glycan_module_cor <- cor(dplyr::select(glycan_select_scores,-grey), method = "spearman")
corcol <-define_colors(as.vector(glycan_module_cor), breaks = c(-1,0,1), quant_colors = c("blue", "white", "red"))

cor_signed <- (glycan_module_cor+1)/2
cor_hc <- fastcluster::hclust(d=as.dist(1-cor_signed))
      
anno_hm <- rowAnnotation(
  site=anno_text(glycan_select_module_anno$site),
  chain=anno_text(glycan_select_module_anno$chain),
  anno=anno_text(glycan_select_module_anno$anno),
  show_annotation_name=rep(TRUE,3)
  )
      
modality_df <- data.frame(feat_modality=glycan_select_module_anno$igx %>% na.omit(),
                          stringsAsFactors = FALSE)
modality_hm <- HeatmapAnnotation(df=modality_df,
                                 which = "row",
                                 show_legend = TRUE,
                                 col=map(as_tibble(modality_df), ~define_colors(.x)),
                                 gp = gpar(col = "white", lwd = 0.5))

hm <- ComplexHeatmap::Heatmap(
  matrix=glycan_module_cor,
  name="glycan inter-module pearson",
  col= corcol,
  cluster_rows = as.dendrogram(cor_hc),
  cluster_columns = rev(as.dendrogram(cor_hc)),
  row_names_gp = gpar(fontsize = 12),
  column_dend_side = "top",
  show_column_names = FALSE,
  right_annotation = modality_hm,
  cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(sprintf("%.1f", glycan_module_cor[i, j]), x, y, gp = gpar(fontsize = 10))
  },
  row_labels = glycan_select_module_anno$module_tag %>% na.omit()
)

hm+anno_hm

```



# WGCNA

Objective of WGCNA is not necessarily globally optimal partitioning, but rather detection of meaningful modules. Grey module is allowed to be very heterogeneous. Also, scale-free property is mostly a heuristic to downscale noise via the power transform, rather than based on theoretical arguments from network/graph theory. Ref Peter Langfelder, Bioconductor forum, https://support.bioconductor.org/p/66101/ [24/04/2019].

How to deal with "grey" module?
  
  Current approach:
  - treat as other clusters, but ignore module-specific modularity contribution in calculation of global modularity?
  - Equivalently: treat all grey members as singleton clusters (so that in modularity calculation: delta(C_i,C_j)=0 for every singleton cluster i and every cluster j)

But also included full modularity, treating the grey cluster as a real cluster.

Modularity is calculated based on three similarity matrices:
  - correlation (following Gomez2009, with known limitations cf MacMahon2015).
- adjacency ~ network parameters
- TOM ~ network parameters

## Glycans

### Including derived traits

#### Correlation distribution

```{r}
glycans_qn_fam_adj_cor <- WGCNA::cor(glycans_qn_fam_adj, 
                                     method = "pearson",
                                     use="pairwise.complete.obs")
glycans_raw_cor <- WGCNA::cor(glycans_raw_omfo, 
                              method = "pearson",
                              use="pairwise.complete.obs")
hist(unlist(glycans_qn_fam_adj_cor))
hist(unlist(glycans_raw_cor))
hist(abs(unlist(glycans_qn_fam_adj_cor))^2)
```



#### Power parameter

```{r glycans-scale-free-softpower}

cowplot::plot_grid(
  glycans_scale_free$p_rsq,
  glycans_scale_free$p_k
)
#low rsq with power distribution (scale free topology): because not independent entities but related measurements?
```

#### Network module parameters

```{r glycans-parameter-exploration, include=FALSE}


# object.size(glycan_module_stats) %>% format(units="MB")
# saveRDS(glycan_module_stats, file = glue("{datadir}/glycan_module_stats.RDS"))
# glycan_module_stats <- readRDS(glue("{datadir}/glycan_module_stats.RDS"))
# View(glycan_module_stats$color_overview)

#scale module-wise modularities?
# glycan_module_stats$plots[[6]]
glycan_module_stats$merge_pars_used %>% 
  map_depth(2,as_tibble) %>%
  map(
    ~bind_cols(.x) %>% 
      setnames(paste0("cut", 1:length(.x))) %>% 
      mutate(parameter=rownames(.x[[1]])) %>% 
      dplyr::select(parameter, everything())
  )
# glycan_module_stats$mod_modularities


# calculate total non-grey modularity, based on weights in adj or in TOM, but with module definition always based on TOM
# remark: sometimes adj-based module definition might be better 
#TODO: retrieve ref for remark


```

#### Networks

```{r glycans-parameter-exploration-plots}
for (i in seq_along(glycan_module_stats$plots)){
  draw(glycan_module_stats$plots[[i]], column_title = glue("Network {i}"))
}

```

Modularity:
  
```{r}
glycan_module_stats$modularity_overview %>% 
  unite(col="nw_cut", network_iter, cut_iter) %>% 
  dplyr::select(nw_cut:cor_mod_modularities_incl_grey) %>% 
  tidyr::gather(mod_def, mod_value, -nw_cut) %>% 
  ggplot(aes(x=mod_def, y=mod_value, color=nw_cut, group=nw_cut))+
  geom_line()+
  coord_flip()

glycan_module_stats$modularity_overview %>% 
  # unite(col="nw_cut", network_iter, cut_iter) %>% 
  # dplyr::select(cor_mod_modularities_incl_grey) %>% 
  mutate_at(vars(adj_mod_modularities_no_grey:cor_mod_modularities_incl_grey),
            function(x){(x-min(x))/diff(range(x))}) %>% 
  tidyr::gather(mod_def, scaled_mod_value, adj_mod_modularities_no_grey:cor_mod_modularities_incl_grey) %>%
  dplyr::filter(str_detect(mod_def, "no_grey")) %>% 
  mutate(mod_def = factor(mod_def, levels = c("cor_mod_modularities_no_grey",
                                              "adj_mod_modularities_no_grey",
                                              "TOM_mod_modularities_no_grey"))) %>% 
  ggplot(aes(x=mod_def, y=scaled_mod_value, color=cut_iter, group=cut_iter))+
  geom_line()+
  facet_grid(networktype~softpower+corfnc)+
  coord_flip()
```


In IgA:
  
  - In HYT (IgA - Ser89-126): three blocks with strong positive intra-block and negative inter-block correlations:
  
  - H>3, high S
- H>3, low S
- H < 5

- cluster of "Bisection of diantennary structures" for different sites, and some of the raw glycans contributing to the derived trait at those

In IgG:
  
  - IgGI low fucose, low SA cluster


### Excluding derived traits

#### power parameter

```{r glycans-scale-free-softpower-noderiv}

# View(glycans_scale_free$sft_df)
cowplot::plot_grid(
  glycans_scale_free_noderiv$p_rsq,
  glycans_scale_free_noderiv$p_k
)
#low rsq with power distribution (scale free topology): because not independent entities but related measurements?
```

#### Network module parameters

```{r glycans-parameter-exploration-noderiv, include=FALSE}


#scale module-wise modularities?
# glycan_module_stats$plots[[6]]
glycan_module_stats_noderiv$merge_pars_used%>% 
  map_depth(2,as_tibble) %>%
  map(
    ~bind_cols(.x) %>% 
      setnames(paste0("cut", 1:length(.x))) %>% 
      mutate(parameter=rownames(.x[[1]])) %>% 
      dplyr::select(parameter, everything())
  )
# calculate total non-grey modularity, based on weights in adj or in TOM, but with module definition always based on TOM
# remark: sometimes adj-based module definition might be better 
#TODO: retrieve ref for remark


```


#### Networks

```{r glycans-parameter-exploration-noderiv-plots}
for (i in seq_along(glycan_module_stats_noderiv$plots)){
  draw(glycan_module_stats_noderiv$plots[[i]], column_title = glue("Network {i}"))
}
```

#### Optimal cut
Criteria per cut:
  Distributions (per cut, per module) of modularity and varExplained do not track.

```{r}
mod_varExpl_plots <- modularity_varExpl_plots(glycan_modules_df_noderiv, varExpl_trh = 0.5)
mod_varExpl_plots$p_mod_sum_varexpl_constraint

```


##### Cut selection

View corrsponding heatmap

```{r}

p=glycan_modules_df_noderiv %>% 
  dplyr::filter(
    corfnc=="bicor",
    networktype=="signed_hybrid",
    powers==6,
  ) %>% 
  dplyr::select(plots) %>% 
  unlist()
p

# # feature cluster labels
glycan_modules_df_noderiv %>%
  unnest(cut_stats) %>%
  dplyr::filter(
    corfnc=="bicor",
    networktype=="signed_hybrid",
    powers==6,
    cut==3
  ) %>%
  unnest(feature_stats) %>%
  group_by(colors) %>% 
  mutate(n=n()) %>% 
  View()

```


#### old

Broken down by network parameters -> same patterns

```{r}

glycan_modules_df_noderiv %>% 
  unnest(cut_stats) %>% 
  unnest(module_stats) %>% 
  tidyr::gather(modularity_type, modularity, 
                matches("mod_modularities")) %>% 
  ggplot(data=., mapping=aes(x=modularity,y=varExplained))+
  geom_point()+
  geom_smooth()+
  facet_grid(powers~modularity_type)

glycan_modules_df_noderiv %>% 
  unnest(cut_stats) %>% 
  unnest(module_stats) %>% 
  tidyr::gather(modularity_type, modularity, 
                matches("mod_modularities")) %>% 
  ggplot(data=., mapping=aes(x=modularity,y=varExplained))+
  geom_point()+
  geom_smooth()+
  facet_grid(corfnc~modularity_type)

glycan_modules_df_noderiv %>% 
  unnest(cut_stats) %>% 
  unnest(module_stats) %>% 
  tidyr::gather(modularity_type, modularity, 
                matches("mod_modularities")) %>% 
  ggplot(data=., mapping=aes(x=modularity,y=varExplained))+
  geom_point()+
  geom_smooth()+
  facet_grid(networktype~modularity_type)

```


```{r}
glycan_module_stats_noderiv$modularity_overview %>% 
  unite(col="nw_cut", network_iter, cut_iter) %>% 
  dplyr::select(nw_cut:cor_mod_modularities_incl_grey) %>% 
  tidyr::gather(mod_def, mod_value, -nw_cut) %>% 
  ggplot(aes(x=mod_def, y=mod_value, color=nw_cut, group=nw_cut))+
  geom_line()+
  coord_flip()

glycan_module_stats_noderiv$modularity_overview %>% 
  # unite(col="nw_cut", network_iter, cut_iter) %>% 
  # dplyr::select(cor_mod_modularities_incl_grey) %>% 
  mutate_at(vars(adj_mod_modularities_no_grey:cor_mod_modularities_incl_grey),
            function(x){(x-min(x))/diff(range(x))}) %>% 
  tidyr::gather(mod_def, scaled_mod_value, adj_mod_modularities_no_grey:cor_mod_modularities_incl_grey) %>%
  dplyr::filter(str_detect(mod_def, "no_grey")) %>% 
  ggplot(aes(x=mod_def, y=scaled_mod_value, color=cut_iter, group=cut_iter))+
  geom_line()+
  facet_grid(networktype~softpower+corfnc)+
  coord_flip()
```


Compared to WGCNA including derivered traits:
  
  Optimal set parameter set is the same (network 2, iteration 3), with slightly higher maximal modularity in the network excluding derived traits.

In IgA:
  
  * In HYT (IgA - Ser89-126): *recovered* three blocks with strong positive intra-block and negative inter-block correlations:
  
  + H>3, high S
+ H>3, low S
+ H < 5

* cluster of "Bisection of diantennary structures" for different sites, and some of the raw glycans contributing to the derived trait at those sites: *not retrieved*
  
  In IgG:
  
  * IgGI low fucose, low SA cluster *is recovered*
  
  ### General remarks
  
  Negative correlations can carry important information, especially when the glycans are separated by addition/removal of a single sugar moiety.

To assure this

1) Structure of the glycan would need to be known
2) Biological evidence of the reaction, e.g. compiled in database such as KEGG?
  
  But: sialic acid typically terminal -> excellent candidate

Examples: 
  ENI1H5N4F0S1 and ENI1H5N4F0S2 are strongly negatively correlated.
LAGY1H5N4F0S2 and LAGY1H5N4F0S1 are strongly positively correlated.

Could relate to:
  
  * regulation of reaction: constitutive versus induced?
  * opposite associations with IPs?
  
  Some derived traits might be driven primarily by a single glycan. Example: ENI_A2S and ENI1H5N4F0S2

-> Possible to assess 'relevance' of derived traits by variance decomposition?
  
## Immunophenotypes
  
```{r wgcna-ip-memory-calc}
# https://peterlangfelder.com/2018/11/25/blockwise-network-analysis-of-large-data/

mem_needed <- function(n_feat){
  bytes <-n_feat^2*8*3
  gigabytes <- bytes/2^(10*3)
  gigabytes
}

max_block_size <- function(GB_RAM){
  bytes <- GB_RAM*2^(10*3)
  n_feat <- sqrt(bytes/24)
  n_feat
}

n_ips <- ncol(ips_trans_fam_adj)
n_ips
mem_needed(n_ips)  #32 GB

max_block_size(32)

```


### Optimal cut

Cuts with deep==4 (i.e. uneven cut indices) perform consistently worse on Gomez modularity.

Ranking is very similar (and identical top 3) when either varExplained-thesholding of varExplained-weighting is used.

```{r}
mod_varExpl_plots <- modularity_varExpl_plots(ip_modules_df, varExpl_trh = 0.5)
mod_varExpl_plots$p_mod_sum_varexpl_constraint
mod_varExpl_plots$p_mod_sum_varexpl_weighted

```


### Cut selection

Explore selected cut

```{r}

ip_cut <- ip_modules_df %>% 
  unnest(cut_stats) %>% 
  dplyr::filter(
    corfnc=="spear",
    networktype=="signed_hybrid",
    powers==5,
    cut==8
  )

# feature labels
# ip_cut %>% 
#   unnest(feature_stats) %>% 
#   View()

```

Module size distribution (excluding grey cluster).

```{r}
ip_cut%>% 
  unnest(module_stats) %>% 
  dplyr::filter(colors!="grey") %>% 
  .$n %>% 
  boxplot()

```


#### Explore annotations

Modules are highly homogenous w.r.t. source

```{r}

plot(ecdf(ip_select_module_anno$max_prop))
  
```

number and size of modules per source

```{r}

ip_select_module_anno %>%
  ggplot(aes(x=label, y=module_size))+
  geom_boxplot(varwidth = TRUE)+
  coord_flip()

```

number of modules per (major) source

```{r}

ip_select_module_anno %>% 
  count(label_major) %>% 
  ggplot(aes(x=reorder(label_major,n), y=n))+
  geom_bar(stat = "identity")+
  coord_flip()

```

compare module-source distribution with feature-source distribution. 

(here the source of a module is the most frequent source in the module ('label_major'),
which is an appropriate approximation because modules are typically highly homogeneous
wrt source, see above)

```{r}

left_join(
  count(ip_select_module_anno, label_major) %>% rename("composite_lin_source"="label_major"),
  dplyr::filter(ip_anno, set_name %in% names(ips_trans_fam_adj)) %>% 
    dplyr::count(composite_lin_source),
  by=c("composite_lin_source"),
  suffix=c("_module", "_feature")
) %>% 
  arrange(n_module) %>% 
  mutate(composite_lin_source=factor(composite_lin_source,composite_lin_source)) %>% 
  gather(dataset, n, -composite_lin_source) %>% 
  ggplot(aes(x=reorder(composite_lin_source,n), y=n))+
  geom_bar(stat = "identity")+
  coord_flip()+
  facet_wrap(~dataset, scales="free_x")

```



# Cross-Modality correlations

## Glycan modules - IPs

Batch effects of lesser relevance.

### Number of significantly associated features per feature group ~ FDR threshold


```{r n-sign-ass-features-FDR}

loadd(igx_mod_ips_n_associated_m)

igx_mod_ips_n_associated_m %>% 
  ungroup() %>%
  dplyr::rename("feature_group"="omic") %>% 
  mutate(
    feature_group = dplyr::recode(
      feature_group,
      IgA="IgA_module",
      IgG="IgG_module",
      modules="IgX_module")
  ) %>% 
  ggplot(aes(x=fdr_threshold, y=prop, colour=feature_group))+
  geom_line()+
  theme_bw()+
  # facet_zoom(x = fdr_threshold < 0.25)
  facet_zoom(xlim = c(0,0.2))+
  xlab("q value")+
  ylab("proportion of features with association < FDR")+
  geom_vline(xintercept=0.05, color="grey")

```


### IP subgroup enrichment

#### Among FDR < 0.05 (dichotomous)

Overall

```{r}

glycan_modules_cor_ips_anno %>% 
  # dplyr::mutate(storey_sign = factor(ifelse(p.val_storey<0.05, "ip_sign", "ip_nonsign"))) %>% 
  group_by(composite_lin_source) %>% 
  summarise(
    n_tests_source = n(),
    n_tests = nrow(glycan_modules_cor_ips_anno),
    n_sign_ip = sum(p.val_storey<0.05),
    n_sign_test = sum(glycan_modules_cor_ips_anno$p.val_storey<0.05),
  ) %>%
  ungroup %>% 
  mutate(p_hyper = ifelse(
    n_sign_ip>0,
    phyper(q=n_sign_ip, m=n_tests_source,
                     n=n_tests-n_tests_source, k=n_sign_test,
                     lower.tail = FALSE),
    1)
  ) %>% 
  View()
  
```

By IgX

```{r}

glycan_modules_cor_ips_anno %>% 
  # dplyr::mutate(storey_sign = factor(ifelse(p.val_storey<0.05, "ip_sign", "ip_nonsign"))) %>% 
  group_by(composite_lin_source, igx) %>% 
  summarise(
    n_tests_source = n(),
    n_sign_ip = sum(p.val_storey<0.05),
  ) %>%
  group_by(igx) %>% 
  mutate(
    n_tests = sum(n_tests_source),
    n_sign_test = sum(n_sign_ip),
  ) %>% 
  ungroup() %>% 
  mutate(p_hyper = ifelse(
    n_sign_ip>0,
    phyper(q=n_sign_ip, m=n_tests_source,
                     n=n_tests-n_tests_source, k=n_sign_test,
                     lower.tail = FALSE),
    1)
  ) %>% 
  arrange(p_hyper) %>% 
  View()
  
```


#### Running (taking association rank into account)

Overall:
  
```{r ip-lin_source-enrich_cont}



p_lin_source_enrich <- igx_mod_ips_lin_source_props_df %>%
  tidyr::gather(lin_source, prop, -one_of("top","weighting")) %>%
  ggplot(aes(x=top, y=prop, color=lin_source, group=lin_source))+
  geom_line()+
  scale_x_log10()+
  facet_wrap(weighting~.)
# ggsave(p_lin_source_enrich, filename = glue("{figdir}/p_lin_source_enrich.png"),
#        width = 6, height = 6, dpi = 600)

p_lin_source_enrich

```

By IgX:
  
```{r lin_source-enrich-by-IgX_cont}

# plot
p_lin_source_IgX_enrich <-  igx_mod_ips_lin_source_IgX_props_df %>%
  tidyr::gather(lin_source, prop, -one_of(c("top", "IgX"))) %>%
  ggplot(aes(x=top, y=prop, color=lin_source, group=lin_source))+
  geom_line()+
  scale_x_log10()+
  facet_grid(IgX~.)
# ggsave(p_lin_source_IgX_enrich, filename = glue("{figdir}/p_lin_source_IgX_enrich.png"),
#        width = 6, height = 6, dpi = 600)

p_lin_source_IgX_enrich

```

## Glycan modules - IP modules

```{r}

glycan_modules_cor_ips_modules_anno %>% 
  View()

```



# Session info

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```


# References

















