---
title: "Immunophenotype CSFs over 100%"
author: "Simon Couvreur"
date: "24/05/2019"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    collapsed: false
    code_folding: hide
    number_sections: true
    self_contained: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(drake)
library(tidyverse)
library(ggplot2)
library(data.table)
library(EnvStats)
library(knitr)
library(kableExtra)
```


```{r}
#load data

loadd(ips_raw_omfo)
loadd(SPEL_omfo)
loadd(ip_anno_dt)
ips_raw_omfo_CSF <- ips_raw_omfo[,-(1:2)] %>% 
  .[, .SD,  .SDcols=setdiff(names(.), SPEL_omfo)]
```

# Data

Data is :

1. intersected (sample-wise) with glycans
2. filtered for 
    a. zero-variance
    b. feature-wise missingness
    c. final_qc_max=="OK"
3. old_p6 features removed (see Mangino2017 for why)
4. filtered for row(sample)-wise missingness
5. intersected (sample-wise) with filtered glycans

After this, some cell subset frequencies (CSF) are above 100%. The following 
analysis is restricted to CSF (not SPEL features).

# The problem

A total of 286 features have at least one CSF > 100%.

```{r}
over_100 <- ips_raw_omfo_CSF[, lapply(.SD, function(x){sum(x>100, na.rm=TRUE)})]
table(unlist(over_100)>0)
```

# Analysis by feature

Overview of the issues ('value' = CSF value, 'feature' = CSF feature):

  - Most features affected (225) are from plate 3
  
```{r}
n_feat <- over_100 %>% 
  transpose() %>% 
  transmute(n_over_100 = V1,
            set_name = names(over_100)) %>%
  inner_join(ip_anno_dt, by="set_name") %>% 
  dplyr::filter(n_over_100>0) %>% 
  count(source)

n_feat%>% 
  ggplot(aes(x=source, y=n))+
  geom_bar(stat="identity")+
  ylab("Number of affected CSF features")+
  xlab("source of features (i.e plate)")


```
  
  - As a proportion of features per plate-source, 'P5 NKT' is most affected 
  (n=23, 12.3% of features)
  
```{r}
over_100 %>% 
  transpose() %>% 
  transmute(n_over_100 = V1,
            set_name = names(over_100)) %>%
  inner_join(ip_anno_dt, by="set_name") %>% 
  mutate(has_over_100 = n_over_100>0) %>% 
  count(source, has_over_100, .drop = FALSE) %>% 
  group_by(source) %>% 
  mutate(n_source = sum(n)) %>% 
  ungroup() %>% 
  mutate(perc=round(n/n_source*100, digits = 1)) %>% 
  dplyr::select(source, has_over_100, perc, n, n_source) %>% 
  dplyr::filter(has_over_100) %>% 
  ggplot(aes(x=source,y=perc)) +
  geom_bar(stat="identity")+
  ylab("Affected CSF features as percentage of total CSF features")+
  xlab("source of features (i.e plate)")
  
```

  - Typically only one value per feature is affected, 
  but a few features are more strongly affected
      * These are typically from plates 4 & 5 (P4, P5 gd, P5 NKT)

```{r}
over_100 %>% 
  transpose() %>% 
  .[V1>0,] %>% 
  summary()

over_100 %>% 
  melt(measure.vars=1:ncol(.), variable.name="set_name", value.name= "n_over_100") %>% 
  .[n_over_100>0,] %>% 
  setkey("set_name") %>% 
  .[setkey(ip_anno_dt, "set_name"), .(n_over_100, source), nomatch=0] %>% 
  ggplot(aes(x=source, y=n_over_100))+
  geom_boxplot()+
  stat_n_text()+
  ylab("number of values>100 per affected feature")+
  xlab("source of features (i.e plate)")
```
     
  - Consistent with two points above, the proportion of _values_ affected is
  largest for P5 NKT and the gap is even stronger than when expressed as proportion
  of _features_ affected.
  
```{r}
over_100 %>% 
  transpose() %>% 
  transmute(n_over_100 = V1,
            set_name = names(over_100)) %>%
  inner_join(ip_anno_dt, by="set_name") %>% 
  group_by(source) %>% 
  summarise(n_value_over_100 = sum(n_over_100),
            n_value = n()*nrow(ips_raw_omfo_CSF),
            ppm = round(n_value_over_100/n_value*1e6)) %>% 
  ggplot(aes(x=source,y=ppm)) +
  geom_bar(stat="identity")+
  xlab("source of features (i.e plate)")+
  ylab("proportion (in ppm) of CSF values (over all features) affected")
    
```

  - When a value is >100, it is typically just over 100, except for P3
  
```{r}
ips_raw_omfo_CSF %>% 
  melt(measure.vars=1:ncol(.), variable.name="set_name", value.name= "value") %>% 
  .[value>100,] %>% 
  setkey("set_name") %>% 
  .[setkey(ip_anno_dt, "set_name"), nomatch=0] %>% 
  ggplot(aes(x=source, y=value))+
  geom_boxplot()+
  stat_n_text()+
  labs(title = "subset: CSFs over 100%")+
  ylab("CSF 'percent' value")+
  xlab("source of features (i.e plate)")
  
```

# Contribution per sample

Because for P3, there is typically just one value>100% per feature and the
exceedance is consistently large: check whether only few samples are underlying
this.

```{r}
ips_raw_omfo_CSF %>% 
  cbind(data.table(index=1:nrow(.)), .) %>%
  melt(id.vars="index", variable.name="set_name", value.name= "value") %>%
  .[value>100,] %>% 
  setkey("set_name") %>% 
  .[setkey(ip_anno_dt, "set_name"), .(index, set_name, value, source), nomatch=0] %>% 
  group_by(source) %>% 
  summarise(n_samples = n_distinct(index)) %>% 
  inner_join(n_feat, by="source") %>% 
  dplyr::rename("n_feat"="n") %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

Indeed only a few (15) samples are causing the exceedances in P3. Now look
at the contribution per sample.

- typically only one plate affected per sample (as might be expected for experimental artefacts)
- Pareto: few sample-plate combinations cause the largest number of exceedances
- **all large exceedances are caused by measurements of sample IID:53612 in P3 (see second plot)**

```{r, fig.asp=2}
ips_raw_omfo[,-1] %>% 
  .[, .SD,  .SDcols=setdiff(names(.), SPEL_omfo)] %>% 
  melt(id.vars="IID", variable.name="set_name", value.name= "value") %>%
  .[value>100,] %>% 
  setkey("set_name") %>% 
  .[setkey(ip_anno_dt, "set_name"), .(IID, set_name, value, source), nomatch=0] %>% 
  ggplot(aes(x=as.factor(IID), fill=source))+
  geom_bar(position = "stack")+
  labs(title = "subset: CSFs over 100%")+
  xlab("IID")+
  ylab("number of exceedances")+
  coord_flip()
```

```{r, fig.asp=2}
ips_raw_omfo[,-1] %>% 
  .[, .SD,  .SDcols=setdiff(names(.), SPEL_omfo)] %>% 
  melt(id.vars="IID", variable.name="set_name", value.name= "value") %>%
  .[value>100,] %>% 
  setkey("set_name") %>% 
  .[setkey(ip_anno_dt, "set_name"), .(IID, set_name, value, source), nomatch=0] %>% 
  ggplot(aes(x=as.factor(IID),y=value, fill=source, color=source))+
  geom_boxplot()+
  labs(title = "subset: CSFs over 100%")+
  xlab("IID")+
  ylab("CSF 'percent' value")+
  coord_flip()+
  stat_n_text(size=3)
```


# Proposed solution

1. Set all P6 CSFs for sample IID:53612 to NA (because CSFs with lower values might
still be overestimated, even when they are not >100%)
2. Cap all other CSFs at 100%.


